Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-07-22 manuscript/bef_stability/manuscript.Rmd

The effect of species richness and network structure on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effect species richness and  network structure on total biomass, productivity
and stability. However, empirical data are lacking on the link between network
structure and ecosystem properties. We used a dataset of standardized fish
sampling in French stream running from 1995 to
2018. We inferred trophic interactions thanks to recorded body size and diet
literature. We then investigate the direct and indirect effects of stream
size and temperature, species richness and network structure on total biomass and
its temporal stability. We found that species richness and network structure had
a positive relationship with total biomass while they had respectively a
negative and positive effect on temporal stability. We also found that
species richness and network structure are not independent and that environment
had an tremendous effect on total biomass and its temporal stability notably
through its effect on species richness and network structure. We conclude that our
study highlights that the effects of species richness and network structure should be 
assessed conjointly and that it opens exciting perspectives on how environment,
species richness and network structure can jointly determine total biomass in natural
communities.

(210/150 words)

@@

Current biodiversity losses and global changes has the potential to affect
ecosystems in complex ways [@bellard_impacts_2012]. Biodiversity has a
tremendous effect on ecosystem properties because it largely determines both the
amount and the distribution of matter and energy fluxes
within ecosystems [@duffy_why_2009]. Biodiversity thus affects biomass
production [@tilman_productivity_1996; @cardinale_species_2002;
@grace_integrative_2016], biomass repartition [@thebault_relationship_2006] and
biomass stability [@thebault_trophic_2005; @thebault_relationship_2006;
@tilman_biodiversity_2006]. The relationship between biodiversity and
ecosystem properties is a central topic in ecology [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009; @isbell_forest_quantifying_2018] and its documentation is thus
crucial to better anticipate the impacts of global changes.

@@

Species richness promotes higher biomass production in a variety of ecosystems
and organisms. Indeed species richness generally increases primary productivity
and total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Theoretical models also
predict this pattern in plant communities [@tilman_plant_1997;
@loreau_partitioning_2001] and food webs [@schneider_animal_2016]. Resource use
complementarity and selection effects have been proposed to elucidate the
positive effect of species richness on productivity. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
A positive selection effect arises when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
In contrast, both positive [@wu_relationship_2014; @li_relationship_2018],
negative or no effects [@szwagrzyk_above-ground_2007; @jacob_productivity_2010;
@seidel_relationship_2013] of species richness on total biomass have been found
in natural forest. Some authors suggest that the lack of resource
complementarity or environmental harshness might be responsible for an absence or a
negative effect of species richness [@loreau_partitioning_2001; @he_global_2013;
@wu_relationship_2014]. Overall, under what conditions and through what
mechanisms we should observe positive relationships between biomass and species
richness in natural settings is still an open question.

Many studies showed that species richness also increases temporal stability,
i.e. variation through time, often quantified using CV. Such positive
relationships between species richness and temporal stability have been reported
in experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018]. Theoretical works
have clarified which mechanisms generate this positive relationship between
species richness and temporal stability [@loreau_species_2008;
@thibaut_understanding_2013]. Two mechanisms may govern the temporal stability
of communities, (1) species synchrony and (2) species variability. Species
synchrony quantify to what extent the fluctuations of abundances or biomasses
through time are correlated among species. A low synchrony in species
fluctuations, i.e.  asynchrony, is synonymous of compensatory dynamic, where the
decrease in abundance of a species is compensated by the increase of an other.
Asynchrony is generally the main mechanism whereby species richness has a
stabilizing effect on temporal stability [@tilman_biodiversity_2006;
@olivier_independent_nodate].  ^[Colin asked for a word about $cv_{sp}$ and
stability but my opinion is that it is not necessary and that the depth of mecanisms are
currently fairly explained for the introduction]. Such positive relationships
between speciesrichness and temporal are also observed in natural settings, such
as stream fish communities [@franssen_annual_2011], bird and butterfly
communities [@olivier_independent_nodate].

@@

One has to notice that most of the empirical and experimental evidence of the
positive effect of richness on biomass production and stablity through time
comes from studies focusing on one trophic level, generally primary producers or
primary consumers. Natural ecosystems generally include
more than one or two trophic levels and it is not certain that what holds
on one trophic level hold also in multitrophic communities.
Theoretical studies showed that food web properties, such as interaction
distribution and strength, affect the distribution of biomass among adjacent
trophic levels, [@thebault_relationship_2006; @schneider_animal_2016;
@barbier_pyramids_2019]. A
large body of theoretical studies links the propagation of perturbations to the
foodweb structure [@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@neutel_reconciling_2007; @duffy_why_2009; @stouffer_compartmentalization_2011],
or trophic chain [@shanafelt_stability_nodate; @wang_biodiversity_2016;
@barbier_pyramids_2019]. To date, we have few evidence on the relationship
between network structure and total biomass and its temporal stability in
empirical multitrophic communities.

@@

When assessing the relationship between community structure and ecosystem
properties in natural communities, it is crucial 
to consider environmental factors [@loreau_biodiversity_2001;
@he_global_2013; @grace_integrative_2016]. Environment affects organism
performance directly, and thereby, species abundances, total biomass
[@grace_integrative_2016] and temporal stability [@de_boeck_patterns_2017;
@hansen_climate_2013; @bluthgen_land_2016; @oliver_heterogeneous_2010; @franssen_annual_2011;
@friedpetersen_drivers_2020]. For example, greater ecosystem size might increase
the number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009]. @li_relationship_2018 have notably shown that climate
moisture has a positive direct effect on total biomass, but also indirect
through a positive effect on species richness. Taking in consideration
environment also allow for an overall comparison of the strength of the
the effects of environment and community structure on total biomass and its
stability.

@@

Stream fish community constitutes a good model to study the effects of
network structure on total biomass and its temporal stability. Despite that
fish trophic interactions are difficult to measure in nature, a consequent body
of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Especially in stream communities, including fishes,
@brose_predator_2019 found that the prediction of predator prey trophic
interactions based on body size ratio was far better in stream ecosystems than
in other aquatic and te ecosystems. Furthermore, the biotic activity in stream
ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain stream ecosystem
properties. Lastly, most of the diversity-biomass and diversity-stability
relationships, have been assessed in terrestrial ecosystems
while there is a evident need to test such relationships in a variety of
ecosystem types.

We used the ONEMA dataset consisting in standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. 99 sites were sampled
for more than 10 years spanning on the 1995-2018 period. We inferred
1496 interaction networks, based on body size and food diet.  The network
structure was summarized by connectance and mean trophic level. Environment
was characterized by the stream size, altitude, slope, and the temperature, flow
and enrichment. Finally total biomass was measured as median total biomass over
the sampling period while its temporal stability was quantified as the inverse
of the CV.

@@

```{r}
matmet_cap <- paste0(
  "(A) Conceptual model of the hypothesized relationships between environment,
  species richness, network metrics and total biomass / biomass stability. ",
  "Solid and dashed arrows represent respectively direct and indirect effects.
  ",
  "The blue, green and blue/green dashed arrows being respectively indirect
  effects going through species richness, network metrics and through both
  species richness and network metrics.",
  "(B) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. ", "Grey shades highlighted elevalation variation. ",
  "Inner borders draw the limits of the hydrographic basins. ",
  "(C) The metaweb. Each trophic species (N = 412) is represented by a node and
  each color represent a species (S = ). Resource nodes (7) were drawn in grey. ",
  "(D) Subset over network variation through time."
)
```

```{r matmet, fig.cap = matmet_cap}
myload(p_st,temporal_network, dir = mypath("manuscript/bef_stability/figs"))

@@

p_method <- plot_grid(top, bottom, nrow = 2, rel_heights = c(1, 1),
  labels = c("", "C"))
save_plot(
  filename = mypath("manuscript/bef_stability/figs",
    "method_fig.pdf"),
  p_method, nrow = 2, ncol =2,
  base_width = 7/2.54,
  base_asp = 0.3
)
p_method
```

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems
(ONEMA) using electric fishing. Two different standardized protocols were used
for little and big streams (detailed in Supplementary methods). For streams
shallower than (XX m), the sampling takes place over the whole width of the
streams and is done by foot whereas in streams deeper than (XX m), the sampling
was done over the bank by foot or by boat. The sampling season spans from 
late spring to Autumn. During sampling, all the fishes caught were
identified and counted. For each species, either the body size of all fish was
measured, or a representative sample when the number of fish was high (See
supplementary methods for details). The fishes were released at the end
of the sampling. Additionally, the physical characteristics of the river was
recorded during sampling as well as the sampling effort, quantified as the
surface sampled.

@@

We kept only the most frequent sampling protocol in each station and kept
stations that had been sampled ten years or more. The median sampling month was
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$, i.e. there was low heterogeneity of the sampling month
within stations. As the concept of community variability referred to variability
around an equilibrium, which is rarely observed or difficult to characterize in
natural communities, only those stations with no temporal trends in total
biomass were selected (detailled in supplementary methods). However, the results
were robust to the relaxation of this hypothesis (Supplementary results). We
retained 99 stations distributed over seven hydrographic basins (Figure
\@ref(fig:matmet) B).

@@

The body size of the unmeasured individuals was inferred under the assumption
that species body size followed a normal distribution whose the mean and the
standard deviation was quantified from the measured individuals of the sampling
(See supplementary methods). The body mass of each individual was then estimated
using an allometric law ($B_i = 0.01 \times l_i^{3.03}$, $B$ being the body mass
in grams, $l$ the body size in millimeters and $i$ the individual), where the
coefficients $0.01$ and $3.03$ comes from meta-analysis
[@lleonart_removing_2000; @froese_cube_2006]. The species biomass
and total biomass at each sampling was then calculated by summing individual
body masses. To account for the sampling effort, species biomass and total
biomass have been related to the surface sampled, i.e. $gm^{-2}$.
In the further analysis, the total biomass of a station has been defined as the
median of the total biomass over time.

@@

with $CV_{com, i}$ being the Coefficient of Variation of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
species in the station $i$ weighted by the biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$) [@loreau_species_2008]. $i$ being
the station $i$ and $k$ the species $k$.

@@

Trophic interaction network was inferred from body size and ontogenic diet shift.
Body size is a reasonable approximation widely used to infer trophic
interactions [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan. For example, alvins feed on plankton while adults may feed on
algae or fishes. Accordingly, each fish species was divided in nine body size
classes, each of which is trophic species (Figure SXX). A high congruence
between inferred interactions and measured trophic interactions has been found
at trophic species scale but not at the species scale [@jennings_weak_2001].
Additionally to the fish trophic species, we added seven resource nodes present
in ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Figure \@ref(fig:matmet)
C & D). The ontogenic diet database was based on informations from fishbase and
literature. Each fish species had two or three life stages which are delimited by
body size (Table SXX, Figure SXX).

The trophic interactions of fish trophic species was determining with the median
of its body size range, hereafter after midpoint. The midpoint of trophic
species fall into one of the species life stage and a trophic link was set
between the trophic species and the resources present in the life stage. If the
trophic species is in a piscivorous life stage, the fish-fish trophic links was
defined in two steps and based on the body size ratio between predators and
prey. First of all, we defined the range predation window from 3% to 45% of the
midpoint of the trophic species. Then, a trophic link was set between the
piscivorous trophic species and every trophic species whose midpoint was
included in the predation range (Figure SXX). Lastly, the trophic links between resources was
set according to the food diet data.

@@

For each trophic network, corresponding to a particular site and sampling, we
computed two of the most commonly used network metrics to characterize the food
web, namely the connectance and mean trophic level. Connectance describes 
the level of generalism in the network and is calculated as the number of actual
interaction divided by the maximum number of interaction if all species interact
together ($C = L / N^2$, L and N being resp. the number of
links and of trophic species plus resource nodes). The trophic level of the
nodes was computed of the nodes that the predator nodes consumes
to which is added one, obtained recursively by setting the trophic level of
basal nodes to 1. The mean trophic level $\overline{T}$ was computed as the mean trophic level
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, with $T_i$ and $B_i$
being resp. the trophic level and the biomass of the node $i$.

@@

we characterised the sampled sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow as they affect the type of fish that can
establish [@franssen_annual_2011; @rolls_scaling_2018; @tonkin_flow_2018;
@gutierrez-canovas_contrasting_2013] as well as BOD (Biological Oxygen Demand)
as a proxy for eutrophisation. As data from the naide database are not collected
on same site as our fish sampling sites, we performed an interpolation with
spatial stream networks [@isaak_applications_2014; @hoef_ssn:_2014]. The
interpolation procedure is detailed in supplementary methods.

We had one environmental variable value for each site and sampling event, except
for water temperature for which the record began in 2006. Then, we computed the average over
time and the CV over time for water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA)
was performed on the 14 scaled variables. We kept in the analysis the two first
principal components based on eigen values distribution (Fig. SXX).
The two first principal components explained `r tot_var_expl_pca`%
of the variance. A varimax axis rotation was performed to maximize the representation
of the variables by the principal components (Figure SXX).
The two first principal components were used in the analysis, representing respectively
the (1) river size (river width & depth), (2) altitude, slope and decreasing
average temperature. Further description of the variables are detailed in
supplementary information.

@@

Structural Equation Modeling are able to assess multivariate relationships
between variables. It is one promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. We built the biomass
model (Figure \@ref(fig:matmet) A, \@ref(fig:sem-plot) B) by considering that environment
can directly affect species richness and network structure, as well as total
biomass. Species richness directly affects total biomass. Food web structure
was shown to be constrained by species richness @winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010. In turn,
the network structure affects total biomass (Figure
\@ref(fig:matmet) A, \@ref(fig:sem-plot) B). We used the same model structure for
temporal variability (Figure \@ref(fig:sem-plot) A). The structural equation
models used for total biomass and its temporal stability contains respectively 4
and 6 response variables, 35 and 26 direct paths.

@@

Environmental variables were interpolated with the `openStars` and `SSN`
packages. Linear models and SEMs were computed with respectively the `nlme` and
`piecewieseSEM` R packages. The code used for the analysis and the Rmarkdown
used to generate this document are accessible on github.

@@

tmp_bm_indir_table <- map(list(env_on_bm, rich_on_bm),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```
```{r}
caption_indir_stab_sem <- paste0(
  "Indirect and total effect of variables on biomass stability from the
  structural equation model. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable, here
  biomass stability. Indirect effects are considered as one step from the target
  variable. For exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is -0.61. To compute indirect
  effect, only the significant direct effects were taken into account. ",
  "Network structure being the indirect effect that went through both avg trophic
  level and connectance and Species richness_Network structure being the
  indirect effect that went through (i) species richness and from (ii) species
  richness to network structure."
)

caption_indir_bm_sem <- paste0(
  "Indirect and total effect of variables on total biomass from the
  structural equation model. See the table 1 for details about computation and
  meaning." 
)
```

@@

tmp_stab_indir_table <- map(list(env_on_stab, rich_on_stab, com_on_stab),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```

```{r indir-bm}
kable(format_table(tmp_bm_indir_table),
  format = "latex",
  booktabs = T,
  label = paste0("indir-bm"),
  caption = caption_indir_bm_sem
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

>>>>

The effect of species richness and network structure on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effects of species richness and network structure on total biomass, productivity
and stability. However, empirical data on the link between network structure and
ecosystem properties are lacking. We used a dataset of standardized fish
sampling in French stream running from 1995 to 2018. We inferred trophic
interactions thanks to recorded body size and diet literature. We then
investigate the direct and indirect effects of stream size and temperature,
species richness and network structure on total biomass and its temporal
stability. We found that species richness and network structure had a positive
relationship with total biomass while they had respectively a negative and
positive effect on temporal stability. We also found that species richness and
network structure are not independent and that environment had a tremendous
effect on total biomass and its temporal stability notably through its effect on
species richness and network structure. Our study highlights that the effects of
species richness and network structure should be assessed conjointly and that it
opens exciting perspectives on how environment, species richness and network
structure can jointly determine total biomass in natural communities.

(208/150 words)

@@

Current biodiversity losses and global changes has the potential to affect
ecosystems in complex ways [@bellard_impacts_2012]. Biodiversity has a
tremendous effect on ecosystem properties because it largely determines both the
amount and the distribution of matter and energy fluxes
within and among ecosystems [@duffy_why_2009]. Biodiversity thus affects biomass
production [@tilman_productivity_1996; @cardinale_species_2002;
@grace_integrative_2016], biomass repartition [@thebault_relationship_2006] and
biomass stability [@thebault_trophic_2005; @thebault_relationship_2006;
@tilman_biodiversity_2006]. The relationship between biodiversity and
ecosystem properties is a central topic in ecology [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009; @isbell_forest_quantifying_2018] and its documentation is thus
crucial to better anticipate the impacts of global changes.

@@

Species richness promotes higher biomass production in a variety of ecosystems
and organisms. Indeed species richness generally increases primary productivity
and total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Theoretical models also
predict this pattern in plant communities [@tilman_plant_1997;
@loreau_partitioning_2001] and food webs [@schneider_animal_2016]. Resource use
complementarity and selection effects have been proposed to elucidate the
positive effect of species richness on productivity. The complementarity in
resource use, rising through niche differentiation and interspecific
facilitation, increase the total amount of resource used
[@loreau_biodiversity_1998; @cardinale_species_2002; @loreau_partitioning_2001].
A positive selection effect arises when species richness leads to an higher dominance
of the most productive species [@tilman_plant_1997; @loreau_partitioning_2001].
In contrast, positive [@wu_relationship_2015; @li_relationship_2018],
negative or non significant effects of species richness on total biomass have been found
in natural forests [@szwagrzyk_above-ground_2007; @jacob_productivity_2010;
@seidel_relationship_2013]. Some authors suggest that the lack of resource
complementarity or environmental harshness might be responsible for an absence or a
negative effect of species richness [@loreau_partitioning_2001; @he_global_2013;
@wu_relationship_2015]. Overall, under what conditions and through what
mechanisms we should observe positive relationships between biomass and species
richness in natural settings remains an open question.

Many studies showed that species richness also increases temporal stability of
communities,
i.e. variation through time, often quantified by the temporal CV. Such positive
relationships between species richness and temporal stability have been reported
in experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018]. Theoretical works
have clarified which mechanisms generate this positive relationship between
species richness and temporal stability [@loreau_species_2008;
@thibaut_understanding_2013], (1) species synchrony and (2) species variability. Species
synchrony quantify to what extent the fluctuations of abundances or biomasses
through time are correlated among species. A low synchrony in species
fluctuations, i.e.  asynchrony, is synonymous of compensatory dynamic, where the
decrease in abundance of a species is compensated by the increase of an other.
Asynchrony is generally the main mechanism whereby species richness has a
stabilizing effect on temporal stability [@tilman_biodiversity_2006;
@olivier_urbanization_2020]. ^[Colin asked for a word about $cv_{sp}$ and
stability but my opinion is that it is not necessary and that the depth of mecanisms are
currently fairly explained for the introduction.] Such positive relationships
between speciesrichness and temporal are also observed in natural settings, such
as stream fish communities [@franssen_annual_2011], bird and butterfly
communities [@olivier_urbanization_2020].

@@

One has to notice that most of the empirical and experimental evidence of the
positive effect of richness on biomass production and stablity through time
comes from studies focusing on one trophic level, generally primary producers or
primary consumers. Natural ecosystems generally include more than one or two
trophic levels and there is no certainty that the relationship between species
richness and total biomass and its temporal stability for a given trophic level
is representative of this relationship at the food web level. Theoretical
studies showed that food web properties, such as interaction distribution and
strength, affect the distribution of biomass among adjacent trophic levels,
[@thebault_relationship_2006; @schneider_animal_2016; @barbier_pyramids_2019]. A
large body of theoretical studies links the propagation of perturbations to the
foodweb structure [@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@neutel_reconciling_2007; @duffy_why_2009; @stouffer_compartmentalization_2011],
or trophic chain [@shanafelt_stability_2018; @wang_biodiversity_2016;
@barbier_pyramids_2019]. To date, we have few evidence on the relationship
between network structure and total biomass and its temporal stability in
empirical multitrophic communities.

@@

When assessing the relationship between community structure and ecosystem
properties in natural communities, it is crucial to consider environmental
factors [@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
Environment affects organism performance directly, and thereby, species
abundances, total biomass [@grace_integrative_2016] and temporal stability
[@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @franssen_annual_2011; @friedpetersen_drivers_2020].
For example, greater ecosystem size might increase the number of available
resources and thus total biomass [@post_ecosystem_2000; @doi_resource_2009].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the the effects of environment and community
structure on total biomass and its stability.

@@

Stream fish community constitutes a good model to study the effects of
network structure on total biomass and its temporal stability. Despite that
fish trophic interactions are difficult to measure in natural conditions, a consequent body
of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Especially in stream communities, including fishes,
@brose_predator_2019 found that the prediction of predator prey trophic
interactions based on body size ratio was far better in stream ecosystems than
in other aquatic and terrestrial ecosystems. Furthermore, the biotic activity in stream
ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain stream ecosystem
properties. Lastly, most of the diversity-biomass and diversity-stability
relationships, have been assessed in terrestrial ecosystems
while there is a evident need to test such relationships in a variety of
ecosystem types.

We used the ONEMA dataset consisting in standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. 99 sites were sampled
for more than 10 years spanning on the 1995-2018 period. We inferred
1496 interaction networks, based on body size and food diet. The network
structure was summarized by connectance and mean trophic level. Environment
was characterized by stream size, altitude, slope, temperature, flow
and enrichment. Finally total biomass was measured as median total biomass over
the sampling period while its temporal stability was quantified as the inverse
of the CV.

@@

```{r}
matmet_cap <- paste0(
  #"(A) Conceptual model of the hypothesized relationships between environment,
  #species richness, network metrics and total biomass / biomass stability. ",
  #"Solid and dashed arrows represent respectively direct and indirect effects.
  #",
  #"The blue, green and blue/green dashed arrows being respectively indirect
  #effects going through species richness, network metrics and through both
  #species richness and network metrics.",
  "(A) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. ", "Grey shades highlighted elevalation variation. ",
  "Inner borders draw the limits of the hydrographic basins. ",
  "(B) The metaweb. Each trophic species (N = 412) is represented by a node and
  each color represent a species (S = 45). Resource nodes (7) were drawn in grey. ",
  "(C) Subset over network variation through time."
)
```

```{r matmet, fig.cap = matmet_cap, cache = FALSE}
myload(p_st,temporal_network, dir = mypath("manuscript/bef_stability/figs"))

@@

left <- plot_grid(p_st, p_metaweb, ncol = 1, rel_widths = c(1, 1), labels =
  c("A", "B"))
p_test <- plot_grid(left, temporal_network, ncol = 2, labels = c("A", "C"))

p_method <- plot_grid(top, bottom, nrow = 2, rel_heights = c(1, 1),
  labels = c("", "C"))
save_plot(
  filename = mypath("manuscript/bef_stability/figs",
    "method_fig.pdf"),
  p_test, nrow = 2, ncol =2,
  base_width = 7/2.54,
  base_asp = 0.3
)
p_test
```

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in Supplementary material, part 1.1). For streams
shallower than 0.7m across the entire width, the sampling took place over the
whole width of the streams and was done by foot whereas in streams deeper than
0.7m, the sampling was done over the bank by foot or by boat. The sampling
season spanned from late spring to autumn. During sampling, all the fishes
caught were identified and recorded. For each species, either the body size of
all fish was measured, or a representative sample when the number of fish was
high (Figure S1, supplementary information). The fishes were released at the end
of the sampling. Additionally, the physical characteristics of the river was
recorded during sampling as well as the sampling effort, quantified as the
surface sampled.

@@

We kept only the most frequent of the two sampling protocols in each station and kept
stations that had been sampled ten years or more. The median sampling month was
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$, i.e. there was low heterogeneity of the sampling month
within stations. As the concept of community variability referred to variability
around an equilibrium, which is rarely observed or difficult to characterize in
natural communities, only those stations with no temporal trends in total
biomass were selected (detailed in supplementary methods). However, the results
were robust to the relaxation of this hypothesis (Supplementary results). We
retained 99 stations distributed over seven hydrographic basins (Figure
\@ref(fig:matmet) B).

@@

The body size of the unmeasured individuals was inferred under the assumption
that species body size followed a normal distribution whose mean and
standard deviation were quantified from the recorded individuals of the sampling
(See supplementary methods, Figure S1). The body mass of each individual was then estimated
using an allometric law ($B_i = 0.01 \times l_i^{3.03}$, $B$ being the body mass
in grams, $l$ the body size in millimeters and $i$ the individual), where the
coefficients $0.01$ and $3.03$ comes from two meta-analyses
[@lleonart_removing_2000; @froese_cube_2006]. Species biomass
and total biomass at each sampling was then calculated by summing individual
body masses. To account for the sampling effort, species biomass and total
biomass have been related to the surface sampled, i.e. $gm^{-2}$.
In the further analysis, the total biomass of a station has been defined as the
median of the total biomass over time.

@@

with $CV_{com, i}$ being the Coefficient of Variation of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
species in the station $i$ weighted by the biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

Trophic interaction network was inferred from body size and ontogenic diet shift.
Body size is a reasonable approximation widely used to infer trophic
interactions [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan. For example, alvins feed on plankton while adults may feed on
algae or fishes. Accordingly, each fish species was divided in nine body size
classes, each corresponding to a trophic species (Figure S2). A high congruence
between inferred interactions and measured trophic interactions has been found
at trophic species scale but not at the species scale [@jennings_weak_2001].
Additionally to the fish trophic species, we added seven resource nodes present
in ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Figure \@ref(fig:matmet)
C & D). The ontogenic diet database was based on fishbase and literature. Each
fish species had two or three life stages which are delimited by body size
(Figure S2).

The trophic interactions of fish trophic species was determined with the median
of its body size range, hereafter midpoint. The midpoint of trophic species fall
into one of the species life stage and a trophic link was set between the
trophic species and the resources present in the life stage. If the trophic
species is in a piscivorous life stage, the fish-fish trophic links was defined
in two steps and based on the body size ratio between predators and prey
[@mittelbach_ontogeny_1998; @claessen_impact_2002]. First of all, we defined the
range predation window from 3% to 45% of the midpoint of the trophic species.
Then, a trophic link was set between the piscivorous trophic species and every
trophic species whose midpoint was included in the predation range (Figure S2).
Lastly, the trophic links between resources was set according to the food diet
data [@allan_stream_2007; @hart_handbook_2008].

@@

For each trophic network, corresponding to a particular site and sampling, we
computed two of the most commonly used network metrics to characterize the food
web, namely the connectance and mean trophic level. Connectance describes 
the level of generalism in the network and is calculated as the number of actual
interaction divided by the maximum number of interaction if all species interact
together ($C = L / N^2$, L and N being resp. the number of
links and of trophic species plus resource nodes). The trophic level of the
nodes was computed of the nodes that the predator nodes consumes
to which is added one, obtained recursively by setting the trophic level of
basal nodes to 1. The mean trophic level $\overline{T}$ was computed as the mean trophic level
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, with $T_i$ and $B_i$
being respectively the trophic level and the biomass of the node $i$.

@@

We characterised the sampled sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow as they affect the type of fish that 
settling down [@franssen_annual_2011; @rolls_scaling_2018; @tonkin_flow_2018;
@gutierrez-canovas_contrasting_2013] as well as BOD (Biological Oxygen Demand)
as a proxy for eutrophisation. As data from the naiades database are not collected
on the same site as our fish sampling sites, we performed an interpolation with
spatial stream networks [@isaak_applications_2014; @hoef_ssn:_2014]. The
interpolation procedure is detailed in supplementary (part 1.2).

We had one environmental variable value for each site and sampling event, except
for water temperature for which the record began in 2006. Then, the temporal
average and CV values of for water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigen values distribution (Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance. A varimax axis rotation was
performed to maximize the representation of the variables by the principal
components (Figure S4). The two first principal components were used in the
analysis, representing respectively the (1) river size (river width & depth),
(2) altitude, slope and decreasing average temperature. Further description of
the variables are detailed in supplementary information (part 1.3).

@@

Structural Equation Modeling are able to assess multivariate relationships
between variables. It is one promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. We built the biomass
model by considering that environment
can directly affect species richness and network structure, as well as total
biomass (Figure \@ref(fig:matmet) A, \@ref(fig:sem-plot) B). 
Food web structure was shown to be constrained by species richness [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010]. 
We assumed that both species richness and network structure directly affect
total biomass (Figure \@ref(fig:matmet) A, \@ref(fig:sem-plot) B). In addition,
we assumed that species richness would indirectly affect total biomass through
network structure [@schneider_animal_2016].
We used the same model structure for temporal variability (Figure
\@ref(fig:sem-plot) A). The structural equation models used for total biomass
and its temporal stability contains respectively 4 and 6 response variables, 35
and 26 direct paths.

@@

Environmental variables were interpolated with the `openStars` and `SSN`
R packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the `nlme` and `piecewieseSEM` R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and the manuscript are accessible on github. The analysis were
performed with R version 3.6.3 [@r_core_team_r_2020].

@@

tmp_bm_indir_table <- map(list(env_on_bm, rich_on_bm),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```

@@

tmp_stab_indir_table <- map(list(env_on_stab, rich_on_stab, com_on_stab),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```
```{r get-sem-single-effect}
stab_sem_coef <- filter(coef_stab_sem, P.Value < 0.05)
